---
title: "Cyclistic Bike Sharing"
author: "Rishabh Chaudhary"
date: "12-12-2022"
output: html_document
---

### Google Analytics Certificate Case Study 1 - Cyclistic
#### Business Task:
Increasing annual members for Cyclistic, how casual riders can be converted to annual membership?

#### Stakeholders:
* Executives: Decision makers (detail-oriented)\n
* Lily Moreno: Marketing Manager (campaigns & initiatives)\n
* Marketing Analytics team\n

#### Scope of this notebook
* Downloading data from source
* Extracting data and move past 12 months data to separate folder
* Read past 12 months data into dataframe
* Rename all columns to make sure the names are consistent
* Combine all dataframes
* Save combined dataframe as CSV

#### Installing required packages
```{r results='hide'}
install.packages("ggplot2")
install.packages("tidyverse")
install.packages("lubridate")
install.packages("dplyr")
```

#### Loading packages
```{r results='hide'}
library("lubridate")
library("ggplot2")
library("tidyverse")
library("dplyr")
```

#### Downloading trip data from web [source](https://divvy-tripdata.s3.amazonaws.com/index.html)
```{r results='hide', eval=FALSE}
# Creating data source URL links
url_prepend <- "https://divvy-tripdata.s3.amazonaws.com/"
url_append <- "-divvy-tripdata.zip"

dates <- c(seq(as.Date(paste0(as.character(202004), '01'), format='%Y%m%d'), as.Date(paste0(as.character(202211), '01'), format='%Y%m%d'), "months")) %>% 
  str_replace_all('-','') %>% 
  str_sub(0,6)

url_complete <- url_prepend %>% 
  paste0(dates, sep="") %>% 
  paste0(url_append, sep="")

for (url in url_complete) {
  download.file(url, "data//" %>%  paste0(str_replace(str_split(url, "//")[[1]][2], "/", "-")))
}

```
#### Downloading stations-trip data
```{r results='hide', eval=FALSE}
urls <- c("https://divvy-tripdata.s3.amazonaws.com/Divvy_Stations_Trips_2013.zip",
          "https://divvy-tripdata.s3.amazonaws.com/Divvy_Stations_Trips_2014_Q1Q2.zip",
          "https://divvy-tripdata.s3.amazonaws.com/Divvy_Stations_Trips_2014_Q3Q4.zip",
          "https://divvy-tripdata.s3.amazonaws.com/Divvy_Trips_2015-Q1Q2.zip",
          "https://divvy-tripdata.s3.amazonaws.com/Divvy_Trips_2015_Q3Q4.zip",
          "https://divvy-tripdata.s3.amazonaws.com/Divvy_Trips_2016_Q1Q2.zip",
          "https://divvy-tripdata.s3.amazonaws.com/Divvy_Trips_2016_Q3Q4.zip",
          "https://divvy-tripdata.s3.amazonaws.com/Divvy_Trips_2017_Q1Q2.zip",
          "https://divvy-tripdata.s3.amazonaws.com/Divvy_Trips_2017_Q3Q4.zip",
          "https://divvy-tripdata.s3.amazonaws.com/Divvy_Trips_2018_Q1.zip",
          "https://divvy-tripdata.s3.amazonaws.com/Divvy_Trips_2018_Q2.zip",
          "https://divvy-tripdata.s3.amazonaws.com/Divvy_Trips_2018_Q3.zip",
          "https://divvy-tripdata.s3.amazonaws.com/Divvy_Trips_2018_Q4.zip",
          "https://divvy-tripdata.s3.amazonaws.com/Divvy_Trips_2019_Q1.zip",
          "https://divvy-tripdata.s3.amazonaws.com/Divvy_Trips_2019_Q2.zip",
          "https://divvy-tripdata.s3.amazonaws.com/Divvy_Trips_2019_Q3.zip",
          "https://divvy-tripdata.s3.amazonaws.com/Divvy_Trips_2019_Q4.zip",
          "https://divvy-tripdata.s3.amazonaws.com/Divvy_Trips_2020_Q1.zip")

for (url in urls) {
  download.file(url, "data//" %>%  paste0(str_replace(str_split(url, "//")[[1]][2], "/", "-")))
}

```

#### Extract and organise data
* Extracted all data/ folder
* Moved past 12 months data fo analysis to separate folder data/Past_12_Months
* Created two columns in all CSVs by opening in excel (ride_length and day_of_week)


#### Read data for past 12 months
```{r results='hide', message=FALSE, warning=FALSE}
# change working directory to Past_12_Months
setwd("C:/Users/RishabhChaudhary/Downloads/certs/Google Analytics Certificate/Case_Study_1_Cyclistic_Bike_Share/data/Past_12_Months/")

# Creating file names for past 12 months
past_12_months <- c(seq(as.Date(paste0(as.character(202112), '01'), format='%Y%m%d'), as.Date(paste0(as.character(202211), '01'), format='%Y%m%d'), "months")) %>% 
  str_replace_all('-','') %>% 
  str_sub(0,6)

# Reading last 12 months data
for (i in past_12_months) {
  file_name <- i %>% paste("-divvy-tripdata.csv", sep = "")
  
  assign("df_" %>% paste(i, sep=""), read_csv(file_name))
}

```

#### Check if column names are consistent for all
```{r}
# Create all dataframe list
df_list <- list(df_202112, df_202201, df_202202, df_202203, df_202204, df_202205, df_202206, df_202207, df_202208, df_202209, df_202210, df_202211)

# Print column names
for (df in df_list) {
  print(colnames(df))
}

# Reset column names
new_colnames <- list("ride_id", "rideable_type", "started_at", "ended_at", "start_station_name", "start_station_id", "end_station_name", "end_station_id", "member_casual", "ride_length", "day_of_week")

df_list_renamed <- lapply(df_list, setNames, new_colnames)
```


#### Combine 12 months data
```{r}
df_combined_12 <- rbind(df_202112, df_202201, df_202202, df_202203, df_202204, df_202205, df_202206, df_202207, df_202208, df_202209, df_202210, df_202211)
head(df_combined_12)
```

#### Summary of combined data
```{r}
summary(df_combined_12)
```

#### Save combined dataset to a csv file
```{r}
# install the data.table package if needed
if (!require("data.table")) install.packages("data.table")

# load the data.table package
library(data.table)

setwd("C:/Users/RishabhChaudhary/Downloads/certs/Google Analytics Certificate/Case_Study_1_Cyclistic_Bike_Share/data/Past_12_Months/")
#write.csv(df_combined_12, "Combined_12_months.csv", row.names = FALSE)

fwrite(as.data.table(df_combined_12), "Combined_12_months.csv")
```
