---
title: "Cyclistic Bike Sharing"
author: "Rishabh Chaudhary"
date: "15-12-2022"
output: html_document
---

### Google Analytics Certificate Case Study 1 - Cyclistic
#### Business Task:
Increasing annual members for Cyclistic, how casual riders can be converted to annual membership?

#### Stakeholders:
* Executives: Decision makers (detail-oriented)\n
* Lily Moreno: Marketing Manager (campaigns & initiatives)\n
* Marketing Analytics team\n

#### Scope of this notebook
* Downloading data from source
* Extracting data and move past 12 months data to separate folder
* Read past 12 months data into dataframe
* Rename all columns to make sure the names are consistent
* Combine all dataframes
* Save combined dataframe as CSV

#### Read data for past 12 months
```{r}
# load the data.table package
library("data.table")

#library(data.table)
library("lubridate")
library("ggplot2")
library("tidyverse")
library("dplyr")

# change working directory to Past_12_Months
setwd("C:/Users/RishabhChaudhary/Downloads/certs/Google Analytics Certificate/Case_Study_1_Cyclistic_Bike_Share/data/Past_12_Months/")

# Reading last 12 months data
df_past_12 <- as.data.frame(fread("Combined_12_months.csv"))
```

#### Inspect combined data
```{r}
#List of column names
print("Column Names:")
colnames(df_past_12)

#How many rows are in data frame?
print("Number of Records:")
nrow(df_past_12)

#Dimensions of the data frame?
print("Dimensions of dataset:")
dim(df_past_12)

#See the first 6 rows of data frame.  Also tail(all_trips)
print("Sample Records:")
head(df_past_12)

#See list of columns and data types (numeric, character, etc)
print("Datatype of variables (columns):")
str(df_past_12)

#Statistical summary of data. Mainly for numerics
print("Statistical Summary:")
summary(df_past_12)

```

#### Cleaning up and adding data to prepare for analysis

* Check distribution in different categories of member_casual
```{r}
# Home many categories in member_casual
table(df_past_12$member_casual)
```

* Add columns that list the date, month, day, and year of each ride
```{r}
#The default format is yyyy-mm-dd
df_past_12$date <- as.Date(df_past_12$started_at, format = "%m/%d/%Y") 
df_past_12$month <- format(as.Date(df_past_12$date), "%m")
df_past_12$day <- format(as.Date(df_past_12$date), "%d")
df_past_12$year <- format(as.Date(df_past_12$date), "%Y")
df_past_12$day_of_week <- format(as.Date(df_past_12$date), "%A")
```

* Inspect the structure of the columns
```{r}
str(df_past_12)
```

* Check distribution in different categories of start_station_name
```{r}
# Home many categories in start_station_name
table(df_past_12$start_station_name == 'HQ QR')
```

#### Conducting descriptice anlysis

* Descriptive analysis on ride_length
```{r}
# Summary of ride_length
summary(df_past_12$ride_length)

```

* Comparing members and casual users
```{r}
aggregate(df_past_12$ride_length ~ df_past_12$member_casual, FUN = mean)
aggregate(df_past_12$ride_length ~ df_past_12$member_casual, FUN = median)
aggregate(df_past_12$ride_length ~ df_past_12$member_casual, FUN = max)
aggregate(df_past_12$ride_length ~ df_past_12$member_casual, FUN = min)

```

* See the average ride time by each day for members vs casual users
```{r}
# Fix order of week days
df_past_12$day_of_week <- ordered(df_past_12$day_of_week, levels=c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))

aggregate(df_past_12$ride_length ~ df_past_12$member_casual + df_past_12$day_of_week, FUN = mean)
```

* analyze ridership data by type and weekday
```{r}
# creates weekday field using wday()
# groups by usertype and weekday
# calculates the number of rides and average duration
# calculates the average duration
# sorts the data
df_past_12 %>% 
  group_by(member_casual, day_of_week) %>% 
  summarise(number_of_rides = n(), average_duration = mean(ride_length, na.rm = TRUE)) %>% 
  arrange(member_casual, day_of_week)
  								
```

*  Let's visualize the number of rides by rider type
```{r}
df_past_12 %>%
  group_by(member_casual, day_of_week) %>% 
  summarise(number_of_rides = n()
            ,average_duration = mean(ride_length, na.rm=TRUE)) %>% 
  arrange(member_casual, day_of_week)  %>% 
  ggplot(aes(x = day_of_week, y = number_of_rides, fill = member_casual)) +
  geom_col(position = "dodge")

```

* Let's create a visualization for average duration
```{r}
df_past_12 %>% 
  group_by(member_casual, day_of_week) %>% 
  summarise(number_of_rides = n()
            ,average_duration = mean(ride_length, na.rm=TRUE)) %>% 
  arrange(member_casual, day_of_week) %>% 
  ggplot(aes(x = day_of_week, y = average_duration, fill = member_casual)) +
  geom_col(position = "dodge")


```

#### Export summary file for further analysis

* Create a csv file that we will visualize in Excel, Tableau, or my presentation software
```{r}
counts <- aggregate(df_past_12$ride_length ~ df_past_12$member_casual + df_past_12$day_of_week, FUN = mean)

setwd("C:/Users/RishabhChaudhary/Downloads/certs/Google Analytics Certificate/Case_Study_1_Cyclistic_Bike_Share/data/Past_12_Months/")

fwrite(as.data.table(counts), "avg_ride_length.csv")

```

